Stabilization of carbon in mineral soils from mangroves of the Sin\'{u} river delta, Colombia

### 2018-06-27 Reviews/Response.tex

\documentclass[11pt]{bgcletter}
\usepackage{hyperref}

\name{Dr.\ Carlos A. Sierra}
\signature{ \vspace{-2cm} Carlos A. Sierra, PhD}
\email{csierra}
\telephone{6133}
\begin{document}
\begin{letter}{Dr. Takashi Asaeda\\
 Editor in Chief \\ Wetlands Ecology and Management}

\opening{Dear Dr. Aseda}
Thank you very much for your consideration of our manuscript and for the opportunity to submit a revised version. We made several changes to the manuscript based on the reviewers' comments. More prominently, we added two new sections, one describing the statistical methods used in the study, and another discussing the implications of our findings. For reproducibility and clarity in our data analysis methods, we also now provide the data and the code necessary to reproduce all results and figures.

 In the text below, we provide answers ({\color{blue} blue font}) to all reviewers' comments ({\it italics font}).

{\bf Reviewer 1} \\
{\it The manuscript of Volkel et al. strengthens previous findings showing that most soil carbon in mangroves is from local production. This is one of the few manuscripts I have seen for this region in Latin America, so it is regionally important. The conclusions on soil carbon origin were supported by mineralogical analyses and stable isotopes, although the stable isotope analyses was incorrectly performed. The design of the study is well conducted, although the number of samples is relatively low (n= 10 plots). The manuscript is well written. See below my detailed suggestions:}

{\color{blue} We thank the reviewer for recognizing the value of the manuscript. Although the reviewer mentions that this study has weaknesses in terms of the isotopic analysis conducted and its representativeness, we believe this is a misunderstanding and clarify these issues in the more specific comments below. }

\begin{itemize}
\item {\it L9- Is better to extrapolate to 1m, instead of 0.8m, to facility comparisons among studies (see (Kauffman and Donato, 2012)}

{\color{blue} We can easily report the values up to 1 m for basin mangroves, but for fringe mangroves soil depth does not go below 0.8 m. It would be wrong to make such an extrapolation. We therefore provide in the abstract a value up to 1 m for the basin mangroves and only up to 0.8 cm for fringe mangroves.}

\item {\it L107- Brand of EL analyser?}
{\color{blue} Elementar. Information added to the manuscript. }

\item {\it L114- Isotopes are ``values'', not signatures, they are not constant. You have to include in your end member analyses: mangrove material, riverine material, marine material (phytoplankton?). You could plot them as a triangle or analyse with mixing model. See (Fry, 2013)}

{\color{blue} We changed `signatures' to `values' as suggested by the reviewer. Please note that we used reported values for the mangrove material as end member in our analysis. We acknowledge that the isotopic analysis could be improved if more end members were available; however, we also want to point out that we not only used C isotopes to identify source contributions, but also mineralogy analyses of river and marine sediments.}

%However, at this point we cannot add a third end member to the analysis since the sampling campaign did not include mangrove material. We considered the sand samples as end members for the marine material, which help to identify inorganic contributions. We acknowledge that the isotopic analysis could be improved if more end members were available; however, we also want to point out that we not only used C isotopes to identify source contributions, but also mineralogy analyses of river and marine sediments. }

\item {\it L117- where were the samples analysed?}

{\color{blue}At the Max Planck Institute for Biogeochemistry in Jena, Germany. This information was added to the manuscript.}

\item {\it L119- What are the standards used for the isotope analyses? }

{\color{blue} The standards were Ali-j3 and Caf-j3. We added a reference were these standards are reported.}

\item {\it L134- How were the samples analysed? }

{\color{blue} We added a section on data analysis describing basic statistical analyses used in this manuscript.}

\item {\it Results \\
You need to add F, df, and p values to all your comparisons. }

{\color{blue} Done.}

\item {\it L156- Make a section on statistical analyses within the Methodology. }

{\color{blue} Done. We are also providing as supplementary material all code and data to reproduce the statistical analyses and figures reported.}

\item {\it L166- Stable isotope values are ratios compared to a standard, so it is obvious that if the value is close to zero, it means that is similar to the standard. No need to explain it. }

{\color{blue} Sentence removed as suggested.}

\item {\it L172- What about sediment deposited in the basin forest after hurricanes, storm surges? Would that contribute to higher riverine sediment in the basin forest? }

{\color{blue} This is irrelevant for our site. During the entire historical hurricane record, no hurricane have arrived to the Colombian Caribbean coast (see Figure below). This is due to the geographical location of this coast, which is protected by the northern part of the South American continent from the main tracks of Atlantic hurricanes. }

\includegraphics[scale=0.15]{noaa-hurricane-tracker.jpg}

\item {\it Fig 2- It would be more interested to have the mean values for the fringe vs basin vs sand vs river material, instead of values for each of the plots. }

{\color{blue} Figure changed as requested.}

\item {\it Fig 3- Better to call the end members `riverine' instead of `Sinu'. Need to include other end members to the analyses such as mangrove material and phytoplankton (maybe some published values ?)}

{\color{blue} We made a number of changes in this figure to improve clarity, including the issue suggested by the reviewer. Published values for other end members are provided in the text.}

\item {\it L221- This is not really interesting or relevant to this paper}

{\color{blue} Although it may not be very interesting, we consider it is relevant to report that the measured values for the sand end member are similar to those of the standard.}

\item {\it L223- And decomposition}

{\color{blue} We changed `heterotrophic consumption by microorganisms' by `decomposition' as suggested.}

\item {\it L259- $\delta$13C is not produced by anything, is just a ratio between the heavy (13C) and the light (12C) isotope. }

{\color{blue} We change `producer' by `contributor' to address this comment.}

\item {\it L266- Need to include all the relevant end members}

{\color{blue} As mentioned before, we used three end-members in our analysis. We also acknowledge that end members for phytoplankton are missing, but our analysis is complemented by the analysis of mineral composition of the sediments. }

\item {\it L357- I am not sure this is the conclusion of the manuscript. Although most of the C seems to be derived from mangroves, basin forests seem to have a higher contribution of other source, either riverine or marine? You could use the N:C values to clarify sources. }

{\color{blue} We found that soils in basin mangroves are dominated by riverine sediments; however, carbon contents in the river samples were close to zero. Instead, we found similar mineralogy among both sediments, but much larger C storage in the basin mangroves. It is well stablished that C has a large affinity to sorb in mineral surfaces of sediments, and our analysis is indicative that the in-situ derived carbon has been stabilized in this mineral surfaces. We believe this is strong evidence to reach this conclusion. }

\item {\it L367- I did not understand how the mineralogical composition facilitates dissolved carbon absorption. Regardless, it doesn't seem to matter much if most of the material is from mangrove origin as it has been shown in previous studies}

{\color{blue} The mineral composition of the sediments determines the surface area available for dissolved organic molecules to sorb. This is well known for different soils and mineral surfaces. To make this more clear, we added a number of additional references and explain this process with more detail in the text. }

\end{itemize}

{\bf Reviewer 2} \\
{\it This paper aims to study source of carbon in a mangrove-wetland complex in Colombia by exploring sediment characteristics. Although a potentially important and interesting study, it appears authors/researchers were not able to gather evidence that might be sufficient to draw the conclusions they presented. This reviewer believes that after critical revision, adding significant new information and providing additional analysis/interpretation their work could be a good addition to the literature, but in its present form, it is not fit for publication.}

{\color{blue} We thank the reviewer for recognizing the importance of this work as well as providing critical comments. We acknowledge that some information was missing in the first version of the manuscript, and we made an effort to include missing details.}

{\it This MS could benefit immensely with a clear layout of sampling plan - sampling stations/plots, mangrove types, distribution and general habitat characteristics, elaborating the logic behind the selection of sampling plots. Attributes such as water depth, tidal range, salinity, proximity to the river mouth, fresh water discharge (rainfall patterns) and tidal mixing etc should be provided. Size of estuarine complex, and annual rainfall etc can be included in study area description.}

{\color{blue} An important detail that we forgot to mention in the submitted version is that this study was part of a much larger project in which a number of plots were stablished previously to estimate carbon stocks in the area. We did not selected the plots ourselves, but rather sampled the set of existing plots. Criteria for plot selection are described in other publications and we refer to them in the new version of the manuscript. }

{\it Also, Figure 1 (map of study location) indicates that river Sinu flows almost 10-12 km west of the sampling region. There is not clear suggestion in the figure or text that samples sites receive waters from River Sinu. This reviewer is somewhat perplexed and unable to understand the hypothesis and choice of the study design. Perhaps some clarification and effort in explaining this via text and figures would be beneficial for the readers.}

{\color{blue} We acknowledge that this point may not have been well explained in the area description section of the previous version of the manuscript. The Sin\'u River changed its course in 1938, and before this year it discharged in the study area. This is well documented in Serrano (2004), and we provide now a better explanation in the description of the study site.}

{\it Some important information is missing in the laboratory analysis section. For this entire section there is no reference/no citation. This reviewer encourages authors to cite appropriate methods that were followed for analysis, and include all necessary and important details.}

{\color{blue} We added necessary citations regarding the different methods used.}

{\it For calculating carbon storages in the sampled sites, bulk density of only one fringe site (P21) and one basin site (C4) was used for all plots. There is a large amount of spatial variability in soils even at small resolution and it is important to use site specific bulk density to calculate site specific storages. This reviewer has a serious reservation against this methodology.}

{\color{blue} Please note the extreme differences in bulk density between the two mangrove types reported in Table 1. While basin mangroves have a bulk density of around 1.1, fringe mangroves are around 0.1 g cm$^{-3}$. We obtained similar results in a previous study that included more replicates (Bolivar 2015), and the numbers reported here simply confirm previous measurements. It is unlikely that these differences are due to spatial variability within mangrove types. In fact, these differences in bulk density are an inherent characteristic of the different mangrove types. The fringe mangroves are stablished in the old Sin\'u river delta sediments, while the fringe mangroves advance the cost line and establish over coastal sands. }

{\it This methods section does not include any description how data was analyzed. No statistical information is provided. What programs were used. Whether any tests for normalcy were performed? It is a standard practice to include such information in the MS. }

{\color{blue} We added a new section describing the statistical procedures and the software used. In addition, we provide the code and data to reproduce all results.}

{\it This reviewer believes that the discussion section of this MS should be revised with more critical inputs, highlighting the new addition to this field. This reviewer commends authors for tedious field work and many hours of data entry and subsequent analysis, but it becomes difficult to see value of this research unless a more rigorous and through discussion is presented. The researchers could also report the broader implications of their research… i.e. How this understanding would be beneficial and for whom.}

{\color{blue} Thanks for the recommendation. We added a paragraph to the discussion about the broader implications of this study.}

{\it There are multiple incidences of discrepancies in sentence syntax and complex sentence formulation that diminishes readability and clarity of the text. This reviewer has suggested moving around text, and rewriting sentences to impart clarity. Such corrections will improve paper's readability and should be relatively straight-forward to fix. }

{\color{blue} Thanks for pointing this out. We edited the text in several parts to reduce sentence complexity and improve clarity.}

\vspace{2em}
{\bf Comments by reviewer 2 on pdf version of the article} \\
The pdf version of the article provided by the reviewer contained a large number of highlighted text, comments, and strikethrough text. We provided below answers to the comments, and the strikethrough text was removed from the manuscript. However we did not performed any action on the highlighted text since there was no indication by the reviewer on the reason the text was highlighted.

{\color{blue} Can you provide an approximate size of this estuarine lagoon complex?} \\
{\it The total area is 5,098 ha. We added this value to the site description section. }

{\color{blue} how about Crystalline iron oxides? } \\
{\it The oxalate extraction dissolves mostly poorly crystalline iron oxides, while the dithionite extraction removes the crystalline Fe oxides. We added this to the text. }

{\color{blue} Figure 4, About 85\% is Aragonite. That also clearly suggests CaCO3 presence. }
{\it Yes, this is correct and we point it out in the main text as well. }

{\color{blue} Ln. 240. Consider rewriting. Doesn't convey meaning clearly} \\
{\it Sentence rewritten as suggested.}

{\color{blue} Ln 303. Looks like there is something missing here. You can provide more information here to build up an explanation for your observation. } \\
{\it This sentence was removed from the latest version of the manuscript because it does not add useful information for the message we want to convey. }

\vspace{2em}
We hope this new version adequately addresses reviewer's comments and it is now suitable for publication.

\closing{Sincerely, \\
 \includegraphics[scale=0.7]{../../../../Documents/Personal/firma.jpg}
 }
 \end{letter}

 \end{document}

### 2018-06-27 Manuscript/carbonStabilizationMangrovesV2.tex
Started working on second round of revisions

\section{Materials and methods} \label{sec:1}
\subsection{Study site}
The study site is located on the northwestern Caribbean coast in Colombia (9$^{\circ}$23'N 75$^{\circ}$52'W), which is part of the southern extreme of Morrosquillo gulf and it is locally known as Cispat\'a bay. The coastal zone is characterized by the Sin\'{u} river delta and a complex estuarine lagoon system that covers approximately 5,098 ha. Extensive wetlands and mangroves dominate this area. The Sin\'{u} river has its origin in the northern part of the western Andean mountains, and between 1938 and 1945 changed its course creating a new delta \citep{suarez2004}. The current estuarine mangrove system was established in the previous river delta. 
Fringe forests are dominated by  \textit{Rhizophora mangle}, basin forests are dominated by \textit{Avicennia germinans}. \textit{Laguncularia racemosa} occurs in both forest types.

\subsection{Field sampling}
Our sampling focused on a set of existing plots previously established to determine the carbon sequestration potential of the mangroves of Cispat\'a bay \citep{Bolivar2015}. We sampled 10 plots up to 1 m in depth from March 12 to 13, 2016, using a soil corer of 7 cm in diameter. We sampled five soil cores at five fringe mangrove sites (plots P21, R1, R4, R5, R6) and at 5  basin mangrove sites (plots P16, C1, C2, C3, C4). In addition to the 10 plots sampled across the mangrove area, 3 nearby sand cores at Nisperal coast were collected, and also 3 riverbed cores of the Sin\'{u} river near the city of Monter\'{i}a, which is located 70 km south from Cispat\'{a} bay (Fig. \ref{fig:1}). Generally, cores without layer changes were divided into sections every 20 cm. We divided conspicuous layer changes at the boundary. Because only the general mineralogical and elemental composition of the sand and river samples was important - we considered them as end members - those samples were not divided into segments. Instead, the whole cores down to 1 m were used as individual samples. \par
Additionally, we selected one fringe and one basin mangrove plot (P21 and C4) to measure bulk density. At each site, a soil pit was dug and samples were collected using sampling rings with a volume of 98.52 cm$^3$. We collected four depth levels at plot P21 (0-20, 20-40, 40-60, 60-80 cm) and five depth levels (+ 80-100 cm) at plot C4. Sampling for bulk density was replicated three times for each depth level. In total, we obtained 60 soil samples out of the mangrove area, 6 end-member samples, and 27 samples for bulk density measurements.
\label{sec:2}

\subsection{Laboratory analyses}
After collection, samples were oven-dried at 70$^{\circ}$C for 5 days. Samples for bulk density measurements were oven dried at 105$^{\circ}$C. We calculated soil bulk density as dry mass divided by fresh volume \mbox{(V = 98.52 cm$^3$)}. Each sample was ground for 3 min at a frequency of 25 Hz using a Retsch MM 400 ball mill.

@@

%\paragraph{$\delta$\textsuperscript{13}C Signatures} 
We used $\updelta^{13}$C values of the sampled material and compared them with the  $^{13}$C/$^{12}$C ratio of the two chosen end-members as indicators for the origin of the carbon. We measured $^{13}$C of all samples using a Finnigan MAT IRMS coupled with an EA 1100 elemental analyzer. Ali-j3 (Acetanilide-Jena3) and Caf-j3 (a caffeine sample from a `Traube synthesis' in large supply) were chosen as internal working standards \citep{werner2001}. All elemental and isotopic analyses were conducted at the Max Planck Institute for Biogeochemistry in Jena, Germany.

%\paragraph{X-ray diffraction analysis}
We conducted X-ray diffraction (XRD) measurements for qualitative and quantitative phase analyses \citep{spiess2009} on 12 representative samples including 4 end-member samples, and 4 samples of basin and fringe mangroves each. Samples were measured 20 min each, from 5 to 60 $^{\circ}2\uptheta$. We determined each mineral phase using the powder diffraction file (PDF) data. We also conducted a Rietveld refinement using Topas (Bruker). To define the type of clays included in the samples, we further measured each before extracting clay fraction from 3 to 70 $^{\circ}2\uptheta$ using ceramic panels. XRD analyses were conducted at the laboratory for Mineralogy and Geochemistry of the Friedrich-Schiller University in Jena, Germany.

@@

We obtained much higher bulk densities for basin than for fringe mangrove soil samples. Average bulk density for fringe mangrove soils was 0.16 g cm$^{-3}$, while for basin mangrove soils it was 1.11 g cm$^{-3}$ (Table \ref{tab:1}). These values confirm previous results on the same area obtained by \citet{Bolivar2015}.

Differences in bulk densities between the two forest types resulted in a completely different distribution of C compared to the previous results based on \%OC alone. A higher carbon storage in the upper layers of basin mangrove soils is clearly outlined (Table \ref{tab:1}). The differences in TOC between basin and fringe mangrove types are significant according to the one-way analysis of variance test ($p$-value $= 0.032$, $F = 4.88$, 48 d.f.). TOC decreased with soil depth within the first four (basin) and three (fringe) layers. There was an increase in TOC for the last measured depth intervals of basin and fringe soils, which can be traced back to higher \%OC values of those layers.

@@

\subsection{Differences in carbon storage}
Lower values of bulk density in fringe mangrove soils can be explained by their high organic matter content (wood residues, leaf debris and roots). Roots claim a large proportion of the soil volume, which strongly decreases bulk density. Instead, basin mangrove samples of Cispat\'{a} bay are characterized by a dense silty composition including a small organic part, which results in higher bulk densities. The increase in bulk density with depth in both mangrove types is likely the result of hydrostatic pressure and time inside the mangrove forest belowground.\par
The higher proportion of organic matter in fringe than in basin soils also leads to higher \%OC values for fringe soils. Basin mangrove soils instead, are characterized by a small portion of organic topsoil. Because \%TN concentrations correlate with \%OC, they also show lower values for basin mangroves. That both end-members (river sediments and sand) are mostly consisting of mineral components is illustrated by their \%OC contents of nearly 0 \%. Because the \%IC concentration of 10 \% of sand samples is similar to the used pure calcium carbonate standard, we concluded that the sand end member is mostly composed of carbonates.

The decrease of \%OC with soil depth is likely the result of the interaction between decomposition, vertical transport of organic matter, and leaching of dissolved carbon in water. Because soil microorganisms utilize nitrogen and bacteria fix nitrogen, the concentration of \%TN also decreases with depth. The fact that concentrations of \%OC and \%TN decrease continuously with soil depth, but still show layers with higher concentrations in depths of 80-100 cm (basin) and 60-80 cm (fringe), may reflect differences in sedimentation rates over time \citep{Bolivar2015}. In Cispat\'{a} bay, silting processes linked to changes in the position of the Sin\'{u} river delta, current sea level rise, flooding regime and fluvial inputs, can generate deep organic layers that may cause the increase of \%OC with depth for both mangrove types \citep{suarez2004}.

@@

Along with the $^{13}$C results, the mineralogical analyses also confirmed that basin mangrove sediments are terrestrially derived. This finding is also reflected in higher metal oxide and hydroxide contents in basin than in fringe mangrove soils. Concentrations of iron and aluminum oxides in basin mangroves were similar to those of the Sin\'{u} river samples, which is a strong indication of fluvial deposition of inland sediments.

@@

\section{Conclusions}
Based on analyses of carbon concentration, bulk density, stable isotopes of carbon, and mineralogy, we found that most carbon stored in soils of Cispat\'a Bay, Colombia, is produced in situ, with little evidence of carbon imported to the area either by fluvial or marine sedimentation. Interior basin mangroves store significantly more carbon in soils than the more ocean exposed fringe mangroves. Sediments transported by the Sin\'u river and deposited in the delta region contain negligible amounts of organic carbon, but the mineralogical composition of these sediments favors the adsorption of dissolved carbon on charged mineral surfaces, which explains the larger levels of C storage in this type of mangroves.

@@

% BibTeX users please use one of
%\bibliographystyle{spbasic}   % basic style, author-year citations
%\bibliography{Reference}

@@

% Non-BibTeX users please use
%\begin{thebibliography}{}
\begin{thebibliography}{24}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{{#1}}
\providecommand{\urlprefix}{URL }
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{DOI~\discretionary{}{}{}#1}\else
  \providecommand{\doi}{DOI~\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi
\providecommand{\eprint}[2][]{\url{#2}}

\bibitem[{Adame and Lovelock(2011)}]{Adame2011}
Adame MF, Lovelock CE (2011) Carbon and nutrient exchange of mangrove forests
  with the coastal ocean. Hydrobiologia 663(1):23--50,
  \doi{10.1007/s10750-010-0554-7},
  \urlprefix\url{https://doi.org/10.1007/s10750-010-0554-7}

\bibitem[{Alongi(2012)}]{Alongi2012}
Alongi DM (2012) Carbon sequestration in mangrove forests. Carbon management
  3(3):313--322

\bibitem[{Bol{\'\i}var(2015)}]{Bolivar2015}
Bol{\'\i}var J (2015) {Soil and root carbon in two types of mangroves in
  Cispat\'a bay (Colombian Caribbean).} Master's thesis, Universidad Nacional
  de Colombia sede Medell\'in

\bibitem[{Bouillon et~al(2008)Bouillon, Connolly, and Lee}]{Bouillion}
Bouillon S, Connolly R, Lee S (2008) Organic matter exchange and cycling in
  mangrove ecosystems: Recent insights from stable isotope studies. Journal of
  Sea Research 59(1):44 -- 58,
  \doi{https://doi.org/10.1016/j.seares.2007.05.001},
  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S1385110107000469},
  mangrove Macrobenthos Special Issue

\bibitem[{Donato et~al(2011)Donato, Kauffman, Murdiyarso, Kurnianto, Stidham,
  and Kanninen}]{Donato2011}
Donato DC, Kauffman JB, Murdiyarso D, Kurnianto S, Stidham M, Kanninen M (2011)
  Mangroves among the most carbon-rich forests in the tropics. Nature Geosci
  4(5):293--297, \urlprefix\url{http://dx.doi.org/10.1038/ngeo1123}

\bibitem[{Farquhar et~al(1989)Farquhar, Ehleringer, , and
  Hubick}]{Farquhar1989}
Farquhar GD, Ehleringer JR, , Hubick KT (1989) Carbon isotope discrimination
  and photosynthesis. Annual Review of Plant Physiology and Plant Molecular
  Biology 40(1):503--537, \doi{10.1146/annurev.pp.40.060189.002443}

\bibitem[{Holmgren(1967)}]{Holmgren1967}
Holmgren GGS (1967) A rapid citrate-dithionite extractable iron procedure1.
  Soil Science Society of America Journal 31(2):210--211,
  \urlprefix\url{http://dx.doi.org/10.2136/sssaj1967.03615995003100020020x}

\bibitem[{Jardine and Siikam\"aki(2014)}]{Jardine2014}
Jardine SL, Siikam\"aki JV (2014) A global predictive model of carbon in
  mangrove soils. Environmental Research Letters 9(10):104,013

\bibitem[{Kaiser and Guggenberger(2000)}]{kaiser2000}
Kaiser K, Guggenberger G (2000) {The role of DOM sorption to mineral surfaces
  in the preservation of organic matter in soils}. Organic Geochemistry
  31(7):711--725

\bibitem[{Lacerda et~al(1995)Lacerda, Ittekkot, and Patchineelam}]{lacerda}
Lacerda L, Ittekkot V, Patchineelam S (1995) Biogeochemistry of mangrove soil
  organic matter: a comparison betweenrhizophoraandavicenniasoils in
  south-eastern brazil. Estuarine, Coastal and Shelf Science 40(6):713 -- 720,
  \doi{https://doi.org/10.1006/ecss.1995.0048},
  \urlprefix\url{http://www.sciencedirect.com/science/article/pii/S0272771485700480}

\bibitem[{McCaffrey et~al(1987)McCaffrey, Lazar, and Holland}]{McCaffrey}
McCaffrey MA, Lazar B, Holland HD (1987) The evaporation path of seawater and
  the coprecipitation of br (super -) and k (super +) with halite. Journal of
  Sedimentary Research 57(5):928--937,
  \doi{10.1306/212F8CAB-2B24-11D7-8648000102C1865D},
  \urlprefix\url{http://jsedres.geoscienceworld.org/content/57/5/928},
  \eprint{http://jsedres.geoscienceworld.org/content/57/5/928.full.pdf}

\bibitem[{Mikutta et~al(2006)Mikutta, Kleber, Torn, and Jahn}]{Mikutta2006}
Mikutta R, Kleber M, Torn MS, Jahn R (2006) Stabilization of soil organic
  matter: Association with minerals or chemical recalcitrance? Biogeochemistry
  77(1):25--56, \doi{10.1007/s10533-005-0712-6},
  \urlprefix\url{https://doi.org/10.1007/s10533-005-0712-6}

\bibitem[{Oades(1988)}]{Oades1988}
Oades JM (1988) The retention of organic matter in soils. Biogeochemistry
  5(1):35--70, \doi{10.1007/BF02180317},
  \urlprefix\url{https://doi.org/10.1007/BF02180317}

\bibitem[{Powers and Veldkamp(2005)}]{powers2005}
Powers JS, Veldkamp E (2005) {Regional variation in soil carbon and $\delta$13C
  in forests and pastures of northeastern Costa Rica}. Biogeochemistry
  72(3):315--336

\bibitem[{Prasad and Ramanathan(2008)}]{prasad2008}
Prasad MBK, Ramanathan A (2008) {Sedimentary nutrient dynamics in a tropical
  estuarine mangrove ecosystem}. {Estuarine, Coastal and Shelf Science}
  80(1):60--66

\bibitem[{Ruttenberg and Goni(1997)}]{ruttenberg1997}
Ruttenberg K, Goni M (1997) {Phosphorus distribution, C: N: P ratios, and
  $\delta$ 13 C oc in arctic, temperate, and tropical coastal sediments: tools
  for characterizing bulk sedimentary organic matter}. Marine Geology
  139(1):123--145

\bibitem[{Schwertmann(1964)}]{Schwertmann}
Schwertmann U (1964) Differenzierung der eisenoxide des bodens durch extraktion
  mit ammoniumoxalat-l{\"o}sung. Zeitschrift f{\"u}r Pflanzenern{\"a}hrung,
  D{\"u}ngung, Bodenkunde 105(3):194--202, \doi{10.1002/jpln.3591050303},
  \urlprefix\url{http://dx.doi.org/10.1002/jpln.3591050303}

\bibitem[{Serrano(2004)}]{suarez2004}
Serrano BE (2004) {The Sin{\'u} river delta on the northwestern Caribbean coast
  of Colombia: bay infilling associated with delta development}. Journal of
  South American Earth Sciences 16(7):623--631

\bibitem[{Spie{\ss} et~al(2009)Spie{\ss}, Teichert, Schwarzer, Behnken, and
  Genzel}]{spiess2009}
Spie{\ss} L, Teichert G, Schwarzer R, Behnken H, Genzel C (2009) {Moderne
  R{\"o}ntgenbeugung: R{\"o}ntgendiffraktometrie f{\"u}r
  Materialwissenschaftler, Physiker und Chemiker}. Springer-Verlag

\bibitem[{Spohn and Giani(2012)}]{Spohn2012}
Spohn M, Giani L (2012) {Carbohydrates, carbon and nitrogen in soils of a
  marine and a brackish marsh as influenced by inundation frequency}.
  Estuarine, Coastal and Shelf Science 107:89--96

>>>>

\section{Materials and methods} \label{sec:1}
\subsection{Study site}
The study site is located on the northwestern Caribbean coast in Colombia (9$^{\circ}$23'N 75$^{\circ}$52'W), which is part of the southern extreme of Morrosquillo gulf and it is locally known as Cispat\'a bay (Fig. \ref{fig:1}). The coastal zone is characterized by the Sin\'{u} river delta and a complex estuarine lagoon system that covers approximately 5,098 ha. Extensive wetlands and mangroves dominate this area. The Sin\'{u} river has its origin in the northern part of the western Andean mountains, and between 1938 and 1945 changed its course creating a new delta \citep{suarez2004}. The current estuarine mangrove system was established in the previous river delta. 
Fringe forests are dominated by  \textit{Rhizophora mangle}, basin forests are dominated by \textit{Avicennia germinans}. \textit{Laguncularia racemosa} occurs in both forest types.

\subsection{Field sampling}
Our sampling focused on a set of existing plots previously established to determine the carbon sequestration potential of the mangroves of Cispat\'a bay \citep{Bolivar2015}. We sampled 10 plots of 500 m$^2$ up to 1 m in depth from March 12 to 13, 2016, using a soil corer of 7 cm in diameter (Eijkelkamp bi-partite gouge auger 04.03, Giesbeek, The Netherlands). We selected five randomly chosen points within each plot to extract soil cores at five fringe mangrove sites (plots P21, R1, R4, R5, R6) and at 5  basin mangrove sites (plots P16, C1, C2, C3, C4). In addition to the 10 plots sampled across the mangrove area, 3 nearby sand cores at Nisperal coast were collected, and also 3 riverbed cores of the Sin\'{u} river near the city of Monter\'{i}a, which is located 70 km south from Cispat\'{a} bay (Fig. \ref{fig:1}). Generally, cores without layer changes were divided into sections every 20 cm to explore differences of carbon concentration with depth. We divided conspicuous layer changes at the boundary, except for the sand and river sediment samples that were analyzed as one single sample for mineralogical and elemental composition.

Additionally, we selected one fringe and one basin mangrove plot (P21 and C4) to measure bulk density. At each site, a soil pit was dug three meters away from one of the corners of the plot and samples were collected using sampling rings with a volume of 98.52 cm$^3$. We collected four depth levels at plot P21 (0-20, 20-40, 40-60, 60-80 cm) and five depth levels (+ 80-100 cm) at plot C4. Sampling for bulk density was replicated three times for each depth level. In total, we obtained 60 soil samples out of the mangrove area, 6 end-member samples, and 27 samples for bulk density measurements.
\label{sec:2}

\subsection{Laboratory analyses}
After collection, samples were oven-dried at 70$^{\circ}$C for 5 days. Samples for bulk density measurements were oven dried at 105$^{\circ}$C. We calculated soil bulk density as dry mass divided by fresh volume \mbox{(V = 98.52 cm$^3$)}. Each sample was ground for 3 min at a frequency of 25 Hz using a ball mill (Retsch MM 400, Haan, Germany).

@@

%\paragraph{$\delta$\textsuperscript{13}C Signatures} 
We used $\updelta^{13}$C values of the sampled material and compared them with the  $^{13}$C/$^{12}$C ratio of the two chosen end-members as indicators for the origin of the carbon \citep{Fry2006}. We measured $^{13}$C of all samples using a Finnigan MAT IRMS coupled with an EA 1100 elemental analyzer. Ali-j3 (Acetanilide-Jena3) and Caf-j3 (a caffeine sample from a `Traube synthesis' in large supply) were chosen as internal working standards \citep{werner2001}. All elemental and isotopic analyses were conducted at the Max Planck Institute for Biogeochemistry in Jena, Germany.

%\paragraph{X-ray diffraction analysis}
We conducted X-ray diffraction (XRD) measurements for qualitative and quantitative phase analyses \citep{spiess2009} on 12 representative samples including 4 end-member samples, and 4 samples of basin and fringe mangroves each. Samples were measured 20 min each, from 5 to 60 $^{\circ}2\uptheta$. We determined each mineral phase using the powder diffraction file (PDF) data. We also conducted a Rietveld refinement using the Topas software for XRD analysis (Bruker Corporation). To define the type of clays included in the samples, we further measured each sample before extracting clay fraction from 3 to 70 $^{\circ}2\uptheta$ using ceramic panels. XRD analyses were conducted at the laboratory for Mineralogy and Geochemistry of the Friedrich-Schiller University in Jena, Germany.

@@

We obtained much higher bulk densities for basin than for fringe mangrove soil samples. Average $\pm$ standard deviation of bulk density for fringe mangrove soils was 0.15 $\pm$ 0.02 g cm$^{-3}$, while for basin mangrove soils it was 1.07 $\pm$ 0.12 g cm$^{-3}$ (Table \ref{tab:1}). These values confirm previous results on the same area obtained by \citet{Bolivar2015}.

Differences in bulk densities between the two forest types resulted in a completely different distribution of C compared to the previous results based on \%OC alone (Table \ref{tab:1}). A higher carbon storage in the upper layers of basin mangrove soils is clearly outlined (Table \ref{tab:1}). The differences in TOC between basin and fringe mangrove types are significant according to the one-way analysis of variance test ($p$-value $= 0.032$, $F = 4.88$, 48 d.f.). TOC decreased with soil depth within the first four (basin) and three (fringe) layers. There was an increase in TOC for the last measured depth intervals of basin and fringe soils, which can be traced back to higher \%OC values of those layers.

@@

\subsection{Differences in carbon storage}
Lower values of bulk density in fringe mangrove soils can be explained by their high organic matter content (wood residues, leaf debris and roots). Roots claim a large proportion of the soil volume, which strongly decreases bulk density \citep{Bolivar2015}. Instead, basin mangrove samples of Cispat\'{a} bay are characterized by a dense silty composition including a small organic part, which results in higher bulk densities. %The increase in bulk density with depth in both mangrove types is likely the result of hydrostatic pressure and time inside the mangrove forest belowground.\par

The higher proportion of organic matter in fringe than in basin soils also leads to higher \%OC values for fringe soils. Basin mangrove soils instead, are characterized by a small portion of organic topsoil. Because \%TN concentrations correlate with \%OC, they also show lower values for basin mangroves. That both end-members (river sediments and sand) are mostly consisting of mineral components is illustrated by their \%OC contents of nearly 0 \%. Because the \%IC concentration of 10 \% of sand samples is similar to the used pure calcium carbonate standard, we concluded that the sand end member is mostly composed of carbonates.

The decrease of \%OC with soil depth is likely the result of the interaction between decomposition, vertical transport of organic matter, and leaching of dissolved carbon in water \citep{Elzein1995, Braakhekke2013, Mathieu2015}. Because soil microorganisms utilize nitrogen and bacteria fix nitrogen, the concentration of \%TN also decreases with depth. The fact that concentrations of \%OC and \%TN decrease continuously with soil depth, but still show layers with higher concentrations in depths of 80-100 cm (basin) and 60-80 cm (fringe), may reflect differences in sedimentation rates over time \citep{Bolivar2015}. In Cispat\'{a} bay, silting processes linked to changes in the position of the Sin\'{u} river delta, current sea level rise, flooding regime and fluvial inputs, can generate deep organic layers that may cause the increase of \%OC with depth for both mangrove types \citep{suarez2004}.

@@

Along with the $^{13}$C results, the mineralogical analyses also confirmed that basin mangrove sediments are terrestrially derived as opposed to marine derived. This finding is also reflected in higher metal oxide and hydroxide contents in basin than in fringe mangrove soils. Concentrations of iron and aluminum oxides in basin mangroves were similar to those of the Sin\'{u} river samples, which is a strong indication of fluvial deposition of inland sediments.

@@

\section{Conclusions}
Based on analyses of carbon concentration, stable isotopes of carbon, and mineralogy, we found that most carbon stored in soils of Cispat\'a Bay, Colombia, is produced in situ, with little evidence of carbon imported to the area either by fluvial or marine sedimentation. Interior basin mangroves store significantly more carbon in soils than the more ocean exposed fringe mangroves. Sediments transported by the Sin\'u river and deposited in the delta region contain negligible amounts of organic carbon, but the mineralogical composition of these sediments favors the adsorption of dissolved carbon on charged mineral surfaces, which explains the larger levels of C storage in this type of mangroves.

@@

% BibTeX users please use one of
\bibliographystyle{spbasic}   % basic style, author-year citations
\bibliography{Reference}

@@

% Non-BibTeX users please use
